---
title: Quality control of the variants found from aligning inserted sequences
output:
   md_document:
     fig_width: 10
     fig_height: 5
     variant: markdown_github
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
```

# Quality control of the variants found from aligning inserted sequences

```{r}
library(ggplot2)
library(dplyr)
vars = read.table('variantsInInsertions.SR.tsv', as.is=TRUE, header=TRUE)
```

## Alignment quality

These calls are based on alignment of the inserted sequences (original insertion vs called insertion that contains variants).
Insertions were first matched by location and size and then aligned.
Even if they were matched we might want to remove situations were the alignment was poor.
What should we use as threshold?
Will it affect the mutation rate experiment that we will use these variants for?

First, a look at the distribution of the alignment quality


```{r}
vars %>% select(id, align.prop.match) %>% unique %>% 
  ggplot(aes(align.prop.match)) + geom_histogram() + theme_bw() + xlab('propotion of matches in alignment') +
  geom_vline(xintercept=.7, linetype=2)
```

We see a long tail. Could we use 0.7 as a minimum alignment identity?
Does it affect the mutation rate?

Using a quick calculation as `number of variants / insertion size`:

```{r}
vars.s = vars %>% group_by(align.prop.match, id, type) %>% mutate(mut.rate=n()/ins.size)

ggplot(vars.s, aes(x=align.prop.match, y=mut.rate)) + geom_point(alpha=.5) +
  xlab('propotion of matches in alignment') + ylab('quick mutation rate') + 
  geom_smooth() + theme_bw() + facet_wrap(~type, scales='free') +
  geom_vline(xintercept=.7, linetype=2)
```

We could filter alignments that where below 70% identity. 
They might not be trustworthy and the quickly estimated mutation rate plateaus around this point so it shouldn't affect the rest of the analysis that much.
